---
title: "p8105_hw2_glj2117"
output: github_document
date: '2022-10-02'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

## Problem 1

Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).

```{r}
transit_data = 
  read_csv("./NYC_Transit_Subway_Entrance_And_Exit_Data.csv", col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
    janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, starts_with("route"), entry, exit_only, vending, entrance_type, ada) %>% 
 mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

```
Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?


Answer the following questions using these data:

How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St 8th Avenue; 125st Broadway; 125st Lenox); the distinct function may be useful here.
How many stations are ADA compliant?
What proportion of station entrances / exits without vending allow entrance?
Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?


## Problem 2

Read and clean the Mr. Trash Wheel sheet:

specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
use reasonable variable names
omit rows that do not include dumpster-specific data
round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)

```{r}
mr_trash_wheel = read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N535") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster))
```


